---
title: "exploratory analysis"
output: github_document
---

The thing we start all files with:

```{r setup, include = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.clour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

sclae_colour_discrete = scale_color_viridis_d()
scale_fill_discrete = scale_fill_viridis_d
```


```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

## `group by`

```{r}
weather_df %>%
  group_by(name, month)
```
Note that there's a line now that says there's a group.

Get rid of it with `ungroup`, can do one variable at a time
```{r}
weather_df %>%
  group_by(name, month) %>% 
ungroup()
```
Can also ungroup by one variable at a time.

## counting things

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(n_obs = n())
```

There are 93 in Jan, 84 in Feb, etc.

Can pipe this right into ggplot!

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(n_obs = n())
```

31 days in Jan, 28 in Feb, etc.^

Can also use `count()`

```{r}
weather_df %>%
  count(month, name = "n_obs")
```


**NEVER** use base R's `table`

```{r, eval=FALSE}
weather_df %>%
  pull(month) %>% 
  table()
```

This isn't a dataframe! 

Other helpful counters

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```


## A digression on 2x2 tables

```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

Case when says given this condition, use this output level.

To make the table:

```{r}
weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```

Use `janitor::tabyl` OR `groupby` and then `summarize` with some `pivoting`.

NOT `table`



