---
title: "ggplot 1"
output: github_document
---
Load in a dataset that we'll use often.

You can also set custom sizes:

```{r}
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig_width = 6,
  fig.asp = .6,
  out_width = "90%"
)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
## Registered S3 method overwritten by 'hoardr':
##   method           from
##   print.cache_info httr
## using cached file: /Users/jeffgoldsmith/Library/Caches/R/noaa_ghcnd/USW00094728.dly
## date created (size, mb): 2020-09-25 14:56:47 (7.519)
## file min/max dates: 1869-01-01 / 2020-09-30
## using cached file: /Users/jeffgoldsmith/Library/Caches/R/noaa_ghcnd/USC00519397.dly
## date created (size, mb): 2020-09-25 14:56:52 (1.699)
## file min/max dates: 1965-01-01 / 2020-03-31
## using cached file: /Users/jeffgoldsmith/Library/Caches/R/noaa_ghcnd/USS0023B17S.dly
## date created (size, mb): 2020-09-25 14:56:54 (0.877)
## file min/max dates: 1999-09-01 / 2020-09-30

weather_df
## # A tibble: 1,095 x 6
##    name           id          date        prcp  tmax  tmin
##    <chr>          <chr>       <date>     <dbl> <dbl> <dbl>
##  1 CentralPark_NY USW00094728 2017-01-01     0   8.9   4.4
##  2 CentralPark_NY USW00094728 2017-01-02    53   5     2.8
##  3 CentralPark_NY USW00094728 2017-01-03   147   6.1   3.9
##  4 CentralPark_NY USW00094728 2017-01-04     0  11.1   1.1
##  5 CentralPark_NY USW00094728 2017-01-05     0   1.1  -2.7
##  6 CentralPark_NY USW00094728 2017-01-06    13   0.6  -3.8
##  7 CentralPark_NY USW00094728 2017-01-07    81  -3.2  -6.6
##  8 CentralPark_NY USW00094728 2017-01-08     0  -3.8  -8.8
##  9 CentralPark_NY USW00094728 2017-01-09     0  -4.9  -9.9
## 10 CentralPark_NY USW00094728 2017-01-10     0   7.8  -6  
## # … with 1,085 more rows
```

## Scatterplot

tmax vs tmin

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax))
```

This does the same thing pretty much:
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```


##Let's fancy it up

Add color? lines? more?

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
## Warning: Removed 15 rows containing missing values (geom_point).
```

Adding smooth curve, making dat points more transparent.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE)
## `geom_smooth()` using method = 'gam' and formula 'y ~ s(x, bs = "cs")'
## Warning: Removed 15 rows containing non-finite values (stat_smooth).
## Warning: Removed 15 rows containing missing values (geom_point).
```

Another way to look at it:

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

One more scatterplot.

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```



## Let's use data manipulation as a part of this

```{r}
weather_df %>%
  filter(name == "CentralPark_NY") %>%
  mutate(
    tmax = tmax * (9/5) + 32,
    tmin = tmin * (9/5) + 32
  ) %>%
  ggplot(aes (x = tmin, y = tmax)) + 
  geom_point()
```


## Stacking geoms

Which geoms do you want?

```{r}
weather_df %>%
    ggplot(aes(x = date, y = tmax, color = name)) + geom_smooth()
```

This next one is like a 2D heat map:

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) + geom_bin2d()
```



## Univariate plots

```{r}
weather_df %>%
  ggplot(aes(x = tmax)) +
  geom_histogram()
```

After ggplot you use a +, before it you use a pipe. 

With color, which isn't great but you can add the `facet_grid` to split them:

```{r}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```


Some other plots...

```{r}
weather_df %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3)
```


Still with `tmax` and `name`

Boxplot
```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

Some poeple like violin plots??

```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin()
```



What about ridge plots? This is what the ridges library is for. You can customize alpha and scale or just not include them here.

```{r}
weather_df %>%
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges(alpha = 0.8, scale = 0.8)
```

## Embedding plots

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3)
```


```{r, fig.width=12}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3)
```






